This is the content of lab 1 
1. Calculate the 10% tax and total payable for the transaction that contains $25 item.
```{r}
price <- 25
tax <- 0.1
total <- (price * tax) + price
print(paste("The total payable amount is :",total))
```
Note: We cannot directly concate string and a value like other programming language so we need to use paste, cat or sprintf.

2. In R, you can ask for user input interactively using the readline() function. This function allows you to prompt the user, take their input as a string, and store it in a variable for further use.
```{r}
name <- readline("Enter your name: ")
cat("Hello, ",name,"!")
```
3. To check datatype
```{r}
var <- "R is fun"
print(is.numeric(var))
print(is.character(var))
print(is.integer(var))
print(is.complex(var))
```
Other functions that can be used to get the data type from a variable are "class" and "typeof"
```{r}
number <- 10
print(class(number))
text <- "I am loving R"
print(typeof(text))
```
Challenges
1.Create a code to show the invoice message to the customer using the appropriate thousands separator. Assumed the customer's name is "Jill", the item is "iPhone 16 Plus 256", the item price is $1,675.4, the tax is 10%, and the expected outcome should be
"Dear Jill, your total invoice is for iPhone 16 Plus 256 including tax is $1,842.94". The value should be taken through user input. Hint: Use readline() function. Make sure to change the data type accordingly.
```{r}
name <- readline("Enter your name: ")
item <- readline("Enter your item: ")
price <- as.numeric(readline("Enter the price of the item: "))
tax <- as.numeric(readline("Enter the tax: "))
total <- (price * tax) + price
cat("Dear ", name,", your total invoice for ", item," including tax is $",format(total, big.mark =","),".", sep = "")
```


