---
title: "Lab7"
output: html_document
date: "2026-09-14"
---

2.1 Leaflet is a powerful R package designed for creating interactive maps. It provides an intuitive way to visualise geographic data and is widely used in data science, spatial analysis and dashboard development.

Ensuring whether the leaflet package is installed or not
```{r}
"leaflet" %in% rownames(installed.packages())
```

Since the above code returned TRUE, we do not need to install the leaflet package again.

After installation, the package must be loaded in R session every time we want to use it

```{r}
library(leaflet)
```

Displaying a basic interactive map of Melbourne
```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = 144.9631, lat = -37.8136, zoom = 10)
```

2.2 Adding Location Data (Markers)
After displaying a basic interactive map, the next step is to enhance it by adding location data which is basically done by using markers that represents specific geographic points on the map

To begin, we first create a simple dataset containing location names along with their corresponding longitude and latitude coordinates:

```{r}
data_points <- data.frame(
  name = c("CBD", "Point A", "Point B"),
  lon = c(144.9431, 144.95, 144.9577),
  lat = c(-37.8136, -37.80, -37.82)
)
```

This dataset consists of three locations in the Melbourne area. Each entry includes:
1. A label (name)
2. A longitude value (lon)
3. A latitude value (Lat)

Once the dataset is prepared we can visualize these locations on the map by adding markers:

```{r}
leaflet(data_points) %>%
  addTiles() %>%
  setView(lng = 144.9631, lat = -37.8136, zoom = 12) %>%
  addMarkers(~lon, ~lat, popup = ~name)
```

2.3 Adding Lines (Polylines)
We can extend the map by adding lines. Lines are useful for representing paths, routes or connections between locations such as transportation routes, delivery paths or movement trajectories

To begin, we define a set of dataset containing a sequence of coordinates that will form a line

```{r}
line_data <- data.frame(
  lng <- c(144.900651, 144.897032, 144.894093),
  lat <- c(-37.794464, -37.796452, -37.792942)
)
```

