---
title: "Lab7"
output: html_document
date: "2026-09-14"
---

2.1 Leaflet is a powerful R package designed for creating interactive maps. It provides an intuitive way to visualise geographic data and is widely used in data science, spatial analysis and dashboard development.

Ensuring whether the leaflet package is installed or not
```{r}
"leaflet" %in% rownames(installed.packages())
```

Since the above code returned TRUE, we do not need to install the leaflet package again.

After installation, the package must be loaded in R session every time we want to use it

```{r}
library(leaflet)
```

Displaying a basic interactive map of Melbourne
```{r}
leaflet() %>%
  addTiles() %>%
  setView(lng = 144.9631, lat = -37.8136, zoom = 10)
```

2.2 Adding Location Data (Markers)
After displaying a basic interactive map, the next step is to enhance it by adding location data which is basically done by using markers that represents specific geographic points on the map

To begin, we first create a simple dataset containing location names along with their corresponding longitude and latitude coordinates:

```{r}
data_points <- data.frame(
  name = c("CBD", "Point A", "Point B"),
  lon = c(144.9431, 144.95, 144.9577),
  lat = c(-37.8136, -37.80, -37.82)
)
```

This dataset consists of three locations in the Melbourne area. Each entry includes:
1. A label (name)
2. A longitude value (lon)
3. A latitude value (Lat)

Once the dataset is prepared we can visualize these locations on the map by adding markers:

```{r}
leaflet(data_points) %>%
  addTiles() %>%
  setView(lng = 144.9631, lat = -37.8136, zoom = 12) %>%
  addMarkers(~lon, ~lat, popup = ~name)
```

2.3 Adding Lines (Polylines)
We can extend the map by adding lines. Lines are useful for representing paths, routes or connections between locations such as transportation routes, delivery paths or movement trajectories

To begin, we define a set of dataset containing a sequence of coordinates that will form a line

```{r}
line_data <- data.frame(
  lng <- c(144.900651, 144.897032, 144.894093),
  lat <- c(-37.794464, -37.796452, -37.792942)
)
```

Visualizing the line on the map:

```{r}
leaflet() %>%
  addTiles() %>%
  addPolylines(
    data = line_data,
    ~lng, ~lat,
    color = "blue",
    weight = 3
) %>%
setView(lng = 144.897, lat = -37.794, zoom = 16)
```


2.4 Adding Areas (Polygons)
In addition to lines, we can also represent areas or regions using polygons. Polygons are commonly used to visualize boundaries such as districts, zones, or coverage areas.
To create a polygon, we define a set of coordinates that form a closed shape (the first and last points should match):

```{r}
polygon_data <- data.frame(
  lng <- c(144.953927, 144.968733, 144.972338, 144.957532),
  lat <- c(-37.812380, -37.808142, -37.816008, -37.820313)
)
```

Then adding polygon to the map

```{r}
leaflet() %>%
  addTiles() %>%
  addPolygons(
    lng = polygon_data$lng,
    lat = polygon_data$lat,
    color = "red",
    fillColor = "red",
    fillOpacity = 0.3
) %>%
setView(lng = 144.9631, lat = -37.8136, zoom = 14)
```

# 3. Thematic Cartography Idioms
Thematic cartography focuses on visualizing spatial data by emphasizing specific attributes or patterns across geographic regions. Thematic maps are used to designed to communicate particular variables such as population, area, density or other statistical measures.

3.1 Loading the dataset

```{r}
library(sf)
url <- "https://ndownloader.figshare.com/files/63164836?private_link=9fbb292910d321311480"

temp_file <- tempfile(fileext = ".geojson")
download.file(url, temp_file, mode = "wb")
geo_data <- st_read(temp_file)
print(geo_data)
```


To verify that the spatial component is correctly loaded, we can visualize the geometry:
```{r}
plot(st_geometry(geo_data))
```

This initial plot provides a basic view of the spatial structure but does not yet incorporate thematic information.
To focus on a specific geographic unit, we can filter and visualize an individual suburb. For example: the following code selects the LGA Hume and displays its boundary:

```{r}
suburb_selected <- geo_data[geo_data$lga_name == "Hume", ]
plot(st_geometry(suburb_selected), col = "lightblue", main =
"Hume")
```

```{r}
suburb_selected <- geo_data[geo_data$lga_name == "Ballarat", ]
plot(st_geometry(suburb_selected), col = "pink", main =
"Ballarat")
```

3.2 Choropleth maps

3.2.1 Introduction
A choropleth map is a type of thematic map where areas (polygons) are colored based on a data value rather than just showing the boundaries. Such maps are commonly used for:
1. Population density
2. Income distribution
3. Sales or business performance by region
4. Public health statistics
By using color intensity, we can quickly identify patterns, trends and outliers geograpthic areas.

3.2.2 Applying choropleth mapping to suburb area
Previously, we visualized spatial data in its basic geometric form (points, lines and areas). However those visualizations only showed where things were located not how they differ.

To demonstrate how choropleth maps work, we will now use sqkm variable from our dataset which represents the area of each LGA in square kilometers.


```{r}
library(ggplot2)
ggplot(data = geo_data) +
  geom_sf(aes(fill = sqkm), color = NA) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(
  title = "Choropleth Map of LGA Area",
  fill = "Area (sqkm)"
)
```

3.3 Symbol maps for Spatial representation

3.3.1 Introduction
While choropleth maps represents data using color with geographic regions, symbols maps use symbols placed at specific location to represent values. These symbols can vary in size, shape or color to encode quantitative or categorical information

3.3.2 Applying symbol mapping to Suburb Area
Since our dataset contains polygon geometries (suburb boundaries), we first need to convert each suburb into a representative point. This is typically done by calculating the centroid of each polygon.

```{r}
library(sf)
geo_points <- st_centroid(geo_data)
```
Each centroid now represents the location of a suburb.
We can now visualize the centroids using ggplot2. The size of each point will represent the sqkm value

```{r}
library(ggplot2)
ggplot() +
  geom_sf(data = geo_data, fill = "grey90", color = "white") +
  geom_sf(data = geo_points, aes(size = sqkm), color = "blue",
alpha = 0.6) +
  theme_minimal() +
  labs(
  title = "Symbol Map of LGA Area",
  size = "Area (sqkm)"
)
```

# 4. Network Data Visualization in R

4.1 Introduction
In addition to spatial data, another important way to represent relationships in data is through network analysis. A network consists of nodes (entities) and edges (connections between entities).
Network visualization is widely used in :
1. Social network analysis (people and relationships)
2. Transportation systems (routes and connections)
3. Communication and interaction data

4.2 Installing and loading required libraries

```{r}
install.packages("igraph")
```
```{r}
library(igraph)
```
4.3 Creating a Simple Network
```{r}
edges <- data.frame(
  from = c("A", "A", "B", "C", "D"),
  to = c("B", "C", "D", "D", "E")
)
```

The dataset describes the relationship such as:
1. A is connected to B and C
2. B is connected to D
3. C is connected to D
4. D is connected to E

Next, we convert this into a network object

```{r}
g <- graph_from_data_frame(edges, directed = FALSE)
```

```{r}
plot(g)
```

